---
title: "Carryover Analysis"
author: "Sibel Kurtoglu"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  word_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, fig.height = 2.5)
```


```{r libraries}
library(readxl)
library(tidyverse)
library(janitor)
library(purrr)
library(openxlsx2)
library(kableExtra)
library(stringr)
library(dplyr)
```



```{r readfiles, eval = FALSE}
# path = "raw_data"
# all_files <- list.files(path, pattern = "*.xlsx$")
# # Use map_df to loop over all files and read each one into a dataframe
# # Combine all dataframes into one, and add .id argument to include the file names
# df_raw <- all_files %>% 
#   set_names() %>%
#   map_df(~ read_excel(file.path(path, .), col_types = c("RT" = "text")), .id = "file_name") %>%
#   clean_names %>%
#   select(file_name, sample_name, compound, type, calculated_amt) %>%
#   mutate(calculated_amt = as.numeric(calculated_amt))
# write.csv(df_raw, "df_raw.csv", row.names = FALSE)
```



```{r, readdfraw}
df_raw <- read.csv("df_raw.csv") 
# 
# 
# df_raw <- df_raw %>%
#   
#   #mutate(date_of_analysis = substr(file_name, start = 1, stop = 6)) %>%
#   mutate(date_of_analysis = str_extract(file_name, "\\d{6}$")) %>%
#   
#   mutate(date_of_analysis = as.Date(date_of_analysis , format = "%y%m%d"))
# 
# print(df_raw)
# 

df_raw <- df_raw %>%
  mutate(date_of_analysis = str_extract(file_name, "[0-9]{6}")) %>%
  mutate(date_of_analysis = as.Date(date_of_analysis , format = "%y%m%d"))




```


```{r removeIS}
# Identify unique compounds
unique_compounds <- unique(df_raw$compound)

# internal standards and real compounds
IS_indices <- str_which(unique_compounds, "-d")
IS_indices2 <- str_which(unique_compounds, ",d")
IS_indices3 <- str_which(unique_compounds, "\\[13\\]")
IS_indices <- c(IS_indices, IS_indices2, IS_indices3)
#unique_compounds[-IS_indices]
real_compounds <- unique_compounds[-IS_indices]
real_compounds <- real_compounds[!is.na(real_compounds)]

df_real <- df_raw %>%
  filter(compound %in% real_compounds)
```


## Background

Background information will go here.

## Methodology
 
## Data Analysis and Processing:


Analysis was performed in the `r version$version.string` and the report generated using RMarkdown. Packages used were openxlsx, readxl, tidyverse, janitor and purrr.
 
The dataset is comprised of UDS 137 instrument reports  from August 29, 2019, to June 28, 2023, from B4-190829 to B332-230628, that is batch 4 on 29-Aug-2019 to batch 332 on 28-Jun-2023. There were `r #length(all_files)` unique runs in the dataset on `r length(real_compounds)` different drugs. Total MRMs monitored in the method, including internal standards is `r length(unique_compounds)`.

True analytes were distinguished from internal standard using regular expression matching for "-d" and "[13]C". 

Data was subsetted by analyte over the `r length(real_compounds)` compound monitored. Data were ordered according to the order they were analyzed in. Consecutive samples that were positive (i.e. above the lowest calibrator) for a drug were identified and it was confirmed that these pairs of consecutive positive samples were i) from the same batch and ii) performed on different patients. 

Concentration data from all paired consecutive positive tests for each analyte were then plotted in a scatterplot with sample N sample N + 1 vs sample N providing a visual representation of the data. This assists in identifying any patterns or anomalies that may represent carryover events. Rapid review of the figures is facilitated by overlay of regression lines. 


## Results

```{r}
# empty dataframe
cal_info <- data.frame(matrix(nrow = 0, ncol = 5))
names(cal_info) <- c("drug", "low_cal_name", "low_cal_value", "high_cal_name", "high_cal_value")
cal_info <- cal_info %>%
  mutate_all(as.character)

# Now you can use the str_detect function

for(i in 1:length(real_compounds)) {
  drug <- real_compounds[i]
  high_cal_name <- df_real %>%
    filter(compound == real_compounds[i]) %>%
    filter(str_detect(sample_name, "^Cal[[:space:]]{1}[0-9]{1}$")) %>%
    pull(sample_name) %>%
    unique() %>%
    tail(1)
  
  high_cal_value <- df_real %>%
    filter(compound == real_compounds[i]) %>%
    filter(sample_name == high_cal_name) %>%
    pull(calculated_amt) %>%
    mean(na.rm = TRUE) %>%
    as.numeric()
  
  low_cal_name <- df_real %>%
    filter(compound == real_compounds[i]) %>%
    filter(!str_detect(sample_name, "^Cal[[:space:]]{1}[0-9]{1}$")) %>%
    pull(sample_name) %>%
    unique()
  
  low_cal_name <- ifelse(is.null(low_cal_name), "No_Low_Cal", low_cal_name)
  
  low_cal_value <- df_real %>%
    filter(compound == real_compounds[i]) %>%
    filter(sample_name == low_cal_name) %>%
    pull(calculated_amt) %>%
    mean(na.rm = TRUE) %>%
    as.numeric()
  
  cal_info[i,] <- c(real_compounds[i], drug,low_cal_name, low_cal_value, high_cal_name, high_cal_value)
}



# Now you can use the replace_na function
cal_info <- cal_info %>%
  mutate(low_cal_value = replace_na(as.numeric(low_cal_value), 0)) %>%
  mutate(high_cal_value = replace_na(as.numeric(high_cal_value), 0))

```

\newpage

#####MY CODE Scatterplot Code

```{r}

# Geriye dönük pozitifleri saklamak için boş bir liste oluşturun
pos_after_pos_list <- list()

# Tüm bileşikler için ggplotlar üretmek için döngü oluşturun
for (i in 1:length(real_compounds)) {

  # Belirli bileşik için veri çerçevesini filtreleyin
  compound_df <- df_real %>%
    filter(compound == real_compounds[i]) %>%
    mutate(casenumber = row_number()) %>%
    relocate(casenumber, .before = sample_name) %>%
    filter(str_detect(sample_name, "^[RE][0-9]{10}")) %>%
    mutate(calculated_amt = replace_na(calculated_amt, 0)) %>%
    arrange(casenumber) %>%
    mutate(next_amt = lead(calculated_amt)) %>%
    
    mutate(next_file = lead(file_name)) %>%
    relocate(next_file, .after = file_name) %>%
    mutate(next_sample = lead(sample_name)) %>% 
    relocate(next_sample, .after = sample_name)
  
  pos_after_pos <- compound_df %>%
    filter((calculated_amt > cal_info$low_cal_value[i]) &  (next_amt > cal_info$low_cal_value[i]) & (file_name == next_file) & (sample_name != next_sample))
  
  # 'zero_after_pos' veri çerçevesinin boyutunu kontrol edin ve boşsa NULL olarak atayın
  if (nrow(pos_after_pos) > 0) {
    pos_after_pos_list[[i]] <- pos_after_pos
  } else {
    pos_after_pos_list[[i]] <- NULL
  }

  # En az 2 satırı olan bir veri çerçevesi varsa ggplot çizmek için kontrol edin
  if (nrow(pos_after_pos) > 2) {
    
    # Ensure the data is numeric and exclude NA values
    compound_df$calculated_amt <- as.numeric(compound_df$calculated_amt)
    compound_df$next_amt <- as.numeric(compound_df$next_amt)
    compound_df <- na.omit(compound_df)
    
    # Correlation test
    correlation_test <- cor.test(compound_df$calculated_amt, compound_df$next_amt)
    correlation_value <- correlation_test$estimate["cor"]
    correlation_CI <- correlation_test$conf.int
   
    # Bileşik için scatterplot oluşturun
    p <- ggplot(pos_after_pos, aes(x = calculated_amt, y = next_amt)) +
      geom_point() +
      geom_smooth(aes(y = next_amt),
                  method = "lm",
                  se = FALSE,
                  color = "red") +
      labs(
        title = paste(
          "Scatterplot of Measured Amount in \nSample (N + 1)  vs Sample N for",
          real_compounds[i]
        ),
        x = "Measured Amount in Sample N",
        y = "Measured Amount in Sample N + 1",
        subtitle = paste("Correlation:", round(correlation_value, 2),
                         "CI:", round(correlation_CI[1], 2), "-", round(correlation_CI[2], 2))
      )   

    # Grafiği göster
    print(p)
    
    # Create the directory if it does not exist
    if(!dir.exists("scatterplot_data")){
      dir.create("scatterplot_data")
    }
    
    # Grafiği Excel dosyasına kaydetmek için
    write_xlsx(pos_after_pos, file.path("scatterplot_data", paste0(real_compounds[i], ".xlsx")))
  }
}

# Listeye bileşik isimlerini atayın
names(pos_after_pos_list) <- real_compounds
print(correlation_test$estimate)




```


\newpage





```{r}

# Create an empty list to store the 'zero after positive' dataframes
zero_after_pos_list <- list()

# Generate ggplots for all compounds
for (i in 1:length(real_compounds)) {
  # Filter the dataframe for a specific compound
  compound_df <- df_raw %>%
    filter(compound == real_compounds[i]) %>%
    mutate(casenumber = row_number()) %>%
    relocate(casenumber, .before = sample_name) %>%
    filter(str_detect(sample_name, "^[RE][0-9]{10}")) %>%
    mutate(calculated_amt = replace_na(calculated_amt, 0)) %>%
    arrange(casenumber) %>%
    mutate(next_amt = lead(calculated_amt)) %>%
    mutate(next_file = lead(file_name)) %>%
    relocate(next_file, .after = file_name) %>%
    mutate(next_sample = lead(sample_name)) %>%
    relocate(next_sample, .after = sample_name)
  
  zero_after_pos <- compound_df %>%
    filter((calculated_amt > cal_info$low_cal_value[i]) &  (next_amt == 0) & (file_name == next_file) & (sample_name != next_sample))
  
  # Check the size of 'zero_after_pos' dataframe and assign it to NULL if it's empty
  if (nrow(zero_after_pos) > 0) {
    zero_after_pos_list[[i]] <- zero_after_pos
  } else {
    zero_after_pos_list[[i]] <- NULL
  }
}
names(zero_after_pos_list) <- real_compounds
```

## historgrams

```{r}
for(i in 1:length(zero_after_pos_list)){
  plot_data <- zero_after_pos_list[[i]] %>%
    filter(
      date_of_analysis >= as.Date("2021-01-01") &
      date_of_analysis <= as.Date("2023-12-31")
    )
  # Check if there's a dataframe with at least 2 rows to draw ggplot
    p_hist <- ggplot(plot_data, aes(x = calculated_amt)) +
      geom_histogram() +
      labs(
        title = paste(
          "Histogram of the Measured Amount in Sample N for",
          real_compounds[i]
        ),
        x = "Measured Amount in Sample N",
        y = "Count",
      )

    # Print the plot
    print(p_hist)
  }

```


# barcharts

```{r}

for(i in 1:length(zero_after_pos_list)) {
  # Create a bar chart for zero after positive events
  plot_data <- zero_after_pos_list[[i]] %>%
    filter(
      date_of_analysis >= as.Date("2021-01-01") &
      date_of_analysis <= as.Date("2023-12-31")
    )
  p_bar <-
    ggplot(plot_data, aes(x = date_of_analysis)) +
    geom_bar() +
    labs(
      title = paste("Number of zero after positive events for", real_compounds[i]),
      x = "Date",
      y = "Number of zero after positive"
    ) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) # Rotate X-axis labels
  
  # Print the bar chart
  print(p_bar)
}

```


\newpage

## Results that are 0 after a Preceeding Positive

Give me a plot of all the results that are 0 after a previously positive result. Make the plots



```{r}
# zero_after_pos_summ veri çerçevesini oluşturmak ve istatistikleri hesaplamak için kullanılan kodlar
zero_after_pos_summ <- data.frame(matrix(nrow = 0, ncol = 9))
names(zero_after_pos_summ) <- c("Drug", "Min", "p05", "1st_Qu.", "Median","3rd_Qu.", "p95", "Max", "Count")

### use mutate_at to make columns 2-7 numeric
zero_after_pos_summ <- zero_after_pos_summ %>% 
  mutate_at(2:9, as.numeric) %>% 
  mutate(Drug = as.character(Drug))

for(i in 1:length(real_compounds)) {
  compound_df <- df_real %>%
    filter(compound == real_compounds[i]) %>%
    mutate(casenumber = row_number()) %>%
    relocate(casenumber, .before = sample_name) %>%
    filter(str_detect(sample_name, "^[RE][0-9]{10}")) %>%
    mutate(calculated_amt = replace_na(calculated_amt, 0)) %>%
    arrange(casenumber) %>%
    mutate(next_amt = lead(calculated_amt)) %>%
    mutate(next_file = lead(file_name)) %>%
    relocate(next_file, .after = file_name) %>%
    mutate(next_sample = lead(sample_name)) %>% 
    mutate(next_case = lead(casenumber)) %>% 
    relocate(next_case, .after = casenumber) %>%
    mutate(is_pos_zero_after = (calculated_amt > cal_info$high_cal_value[i]) & (calculated_amt > next_amt) & (next_amt == 0)) ### code added here ###
  
  zero_after_pos <- compound_df %>%  
    filter(is_pos_zero_after == TRUE)
  
  compound_pos_zero_sum <- compound_df %>% 
    group_by(compound) %>% 
    summarize(positive_zero = sum(is_pos_zero_after, na.rm = TRUE)) ### code added here ###
  
  zero_after_pos <- compound_df %>%
    filter((calculated_amt > cal_info$high_cal_value[i]) & (calculated_amt > next_amt) & (next_amt == 0) & (file_name == next_file) & (sample_name != next_sample))
  
  # Compute summary statistics
  summary_stats <- c(quantile(zero_after_pos$calculated_amt, p = c(0, 0.05, 0.25, 0.5, 0.75, 0.95, 1)), length(zero_after_pos$calculated_amt))
  
  compound_summ <- data.frame(matrix(nrow = 0, ncol = 9)) %>%
    mutate_at(2:9, as.numeric) %>%
    mutate_at(1, as.character)
  names(compound_summ) <- c("Drug", "Min", "p05", "1st_Qu.", "Median","3rd_Qu.", "p95", "Max", "Count")
  compound_summ[1,1] <- c(Drug = real_compounds[i])
  compound_summ[1,2:9] <- c(summary_stats)
  
  # Put drug name in column 1 and summary statistics in columns 2-7
  zero_after_pos_summ <- bind_rows(zero_after_pos_summ, compound_summ)
}
```

```{r}
# Convert zero_after_pos_summ to a kable object
zero_after_pos_summ %>% 
  kbl(booktabs = TRUE, longtable = TRUE, caption = "Summary Statistics for Each Compound", digits = 1) %>%
  kable_styling(latex_options = c("striped"))
```








